\documentclass[10pt]{article}
\usepackage{charter}
\usepackage{textcomp}
\usepackage{stmaryrd}
\usepackage{fullpage}
\usepackage[usenames,dvipsnames,svgnames]{xcolor}
\usepackage{hyperref}
\usepackage[flushleft,neverdecrease,neveradjust]{paralist}
\pagestyle{empty}

\usepackage[T1]{fontenc}

\renewcommand{\labelitemi}{}
\renewcommand{\labelitemii}{}

\hypersetup{
  colorlinks=true,
  urlcolor=Sepia,
  pdfborder= 0 0 0,
  bookmarks=false,
  pdftitle={Siddharth Narayanaswamy - Teaching Statement},
  pdfauthor={Siddharth Narayanaswamy},
  pdfsubject={Teaching Statement}}

\begin{document}

\begin{flushleft}
  \begin{tabular*}{\textwidth}{@{}l@{\extracolsep{\fill}}r@{}}
    \textbf{\huge{Teaching Statement}} & \today
  \end{tabular*}
\end{flushleft}

\begin{flushleft}
  \begin{tabular*}{\textwidth}{@{}l@{\extracolsep{\fill}}l@{}}
    \textbf{\large{Siddharth Narayanaswamy}} & \url{http://www.iffsid.com}
  \end{tabular*}
\end{flushleft}
\vspace*{0.4in}

A significant factor in my preference for academia over industry is the
opportunity to engage and interact with students.
%
The undeniable importance of education in society and the value of sharing the
knowledge we acquire are further factors in strengthening such a preference.
% needs work : transition
\vspace*{1ex}

During the course of my thesis, although I've been funded as a research
assistant through a multitude of grants, I have served as an unofficial
teaching assistant for four offerings of the undergraduate AI course (ECE 473)
and four offerings of the graduate AI course (ECE 570), each such course
offering comprising approximately 20 students.
%
In such a capacity, I held office hours, marked assignments and exams,
interacted with the students via email, and taught the odd lecture.
%
Furthermore, as part of the Purdue Summer Undergraduate Research Fellowships
(SURF) program, I've mentored four undergraduate students, introducing them to,
and guiding them through, the process of conducting research.
%
I also had the opportunity to recruit yet more students in the course of my own
research, orchestrating data-gathering and analysis efforts that resulted in a
publication [{\color{RoyalBlue} C7}] with all the students as co-authors.
% needs work : given me experience in handling students blah blah
\vspace*{1ex}

My research, which involves algorithms, robotics, natural langauge processing,
computer vision, medical imaging, digital signal processing, machine learning,
compilers, optimization, functional programming, programming languages,
distributed systems, and AI, has given me the opportunity to develop a firm
grounding in software, hardware, and analysis.
%
Such a grounding proved to be invaluable in my endeavours thus far; and, I
believe, will continue to be so in the future.
%

% I am happy to teach any course in the undergraduate CS curriculum. A number of
% courses are directly related to my research including machine learning,
% robotics, AI, computer vision, financial engineering, medical imaging, and
% natural language processing. I have extensive hardware and software experience
% having build robots and deployed large-scale systems at Amazon and can teach
% compilers, operating systems, computer architecture, networking, distributed
% systems, programming languages, and functional programming.

\end{document}
